\documentclass{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{array}
\usepackage{fancyhdr}
\usepackage{amssymb}
\usepackage[shortlabels]{enumitem}

\DeclareMathOperator{\R}{\mathbb R}

\pagestyle{fancy}
\fancyhead[L]{Banghao Chi}
\fancyhead[C]{Homework 6}
\fancyhead[R]{12th Apr}

\fancyfoot[C]{\thepage}

\renewcommand{\headrulewidth}{0.5pt}
\renewcommand{\footrulewidth}{0.5pt}

\begin{document}

\section*{Exercise 1}
An interval matrix $A \in \{0, 1\}^{m\times n}$ is such that all 1s in each row occur consecutively. Precisely this means that for all $1 \leq i \leq m$, if $a_{ij} = 1$ and $a_{ik} = 1$ with $j \leq k$, then $a_{ih} = 1$ for all $j \leq h \leq k$. Prove that interval matrices are totally unimodular. \\

\textit{Hint:} Recall that $\det(B) = \det(B')$ if $B'$ is obtained from $B$ by applying the column operation $C_i = C_i + kC_j$ where $j \neq i$. \\

\textbf{Solution:} \\

Let $B$ be any square submatrix of size $n \times n$ from the interval matrix $A$. Due to the interval property, in each row of $B$, the 1s occur consecutively. \\

For each row $i$ of $B$, let $j_i$ be the leftmost column such that $B_{i,j_i} = 1$. By the interval property, if $B_{i,j_i} = 1$ and $B_{i,k} = 1$ with $j_i < k$, then $B_{i,h} = 1$ for all $j_i \leq h \leq k$. \\

For each row $i$, for all columns $k > j_i$ where $B_{i,k} = 1$, we apply the column operation:
\begin{align*}
C_k = C_k - C_{j_i}
\end{align*}

This would replace all 1s in row $i$ (except the leftmost) with 0s. After performing this for all rows, we have a matrix $B'$ where each row has at most one 1. \\

For the determinant of $B'$: \\

Case 1: If $B'$ has a row with no 1s, then $\det(B') = 0$. \\

Case 2: If $B'$ has a column with no 1s, then $\det(B') = 0$. \\

Case 3: Otherwise, each row and each column of $B'$ contains exactly one 1. In this case, $B'$ is a permutation matrix, and the determinant of a permutation matrix is either 1 or -1, depending on the parity of the permutation. \\

Since column operations of the form $C_i = C_i + kC_j$ ($i \neq j$) preserve the determinant, we have:
\begin{align*}
\det(B) = \det(B') \in \{-1, 0, 1\}
\end{align*}

Therefore, every square submatrix of the interval matrix $A$ has determinant in $\{-1, 0, 1\}$, proving that interval matrices are totally unimodular.

\newpage

\section*{Exercise 2}
Consider the bipartite graph with vertices $\{a_1, a_2, \ldots, a_{10}\}$ on one side, vertices $\{b_1, b_2, \ldots, b_{10}\}$ on the other side, and an edge between $a_i$ and $b_j$ if the product $ij$ is a multiple of 6. \\

Find a largest matching in this graph, and show that it cannot be any larger by finding a vertex cover of the same size. \\

\textbf{Solution:} \\



\newpage

\section*{Exercise 3}
A bipartite graph $(X,Y,E)$ has $|X| = |Y| = n$ and is $r$-regular: every vertex (in $X$ or in $Y$) is the endpoint of exactly $r$ edges.

\begin{itemize}
    \item[(a)] Determine $|E|$, the number of edges in the graph.
    \item[(b)] Show that any vertex cover must contain at least $n$ vertices.
    
    This implies that there is a matching of size $n$, which matches every vertex in $X$ to a vertex in $Y$.
\end{itemize}

\textbf{Solution:} \\

\begin{itemize}
    \item[(a)] $|E| = rn$

    \item[(b)] Suppose there exists a vertex cover $C$ with $|C| < n$ vertices.

    Let $C_X = C \cap X$ and $C_Y = C \cap Y$ be the vertices in the cover from sets $X$ and $Y$ respectively.
    So $|C_X| + |C_Y| = |C| < n$.
    
    Since $C$ is a vertex cover, there cannot be any edge between $X - C_X$ and $Y - C_Y$.
    Otherwise, such an edge would not be covered by any vertex in $C$.
    
    The number of vertices not in the cover are:
    $|X - C_X| = n - |C_X|$
    $|Y - C_Y| = n - |C_Y|$
    
    Since $|C_X| + |C_Y| < n$, we have:
    $|C_X| < n - |C_Y|$
    $|X - C_X| > |Y - C_Y|$
    
    Since the graph is $r$-regular, each vertex in $X - C_X$ has exactly $r$ neighbors, all of which must be in $C_Y$ (as no edge can connect to $Y - C_Y$).
    
    So the total number of edges between $X - C_X$ and $C_Y$ is:
    $r \cdot |X - C_X| = r(n - |C_X|)$
    
    But each vertex in $C_Y$ can be incident to at most $r$ edges.
    So the maximum number of edges that can be incident to vertices in $C_Y$ is:
    $r \cdot |C_Y|$
    
    This gives us:
    $r(n - |C_X|) \leq r|C_Y|$
    $n - |C_X| \leq |C_Y|$
    $n \leq |C_X| + |C_Y| = |C|$
    
    This contradicts our assumption that $|C| < n$.
    
    Therefore, any vertex cover must contain at least $n$ vertices.
\end{itemize}

\newpage

\section*{Exercise 4}
Find examples of networks with the following properties:

\begin{itemize}
    \item[(a)] A network with a unique maximum flow, but multiple minimum cuts.
    \item[(b)] A network with multiple maximum flows, but a unique minimum cut.
    \item[(c)] A network with multiple maximum flows and multiple minimum cuts.
\end{itemize}

For each example, describe the maximum flow(s) and the minimum cut(s). \\

\textbf{Solution:} \\



\newpage

\section*{Exercise 5}
Consider the network below, with label $x(y)$ denoting a flow of $x$ and a total capacity of $y$ along an edge. Draw the residual graph, and use it to list all possible augmenting paths. \\

\textbf{Solution:} \\



\end{document}